Verification of concurrent design patterns with data

### 2019-04-18 2018-Arch-pNets/Coordination Camera ready/arch-pNets.tex
newcoord macro et quelques explications

@@

%\paragraph{Definitions.}
%\label{section:pNets}
A pLTS is a labelled transition system with variables occurring inside states,  actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
Variables of each state are pairwise disjoint.
The parameters of actions that appear on pLTS transitions are either input variables, bound when the action occurs, or expressions using the value of other (i.e. output) variables.
We  define the set of parameterised actions a pLTS can use ($a$
ranges over action labels):
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
e_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $e_j^{j \in J}$ 
are expressions.

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
%\begin{itemize}
%\item
$S$ is a set of states; %$\vars(s)$ denotes the set of associated variables;
%\item
$s_0 \in S$ is the initial state;
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
%\item 
$\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
% SB: I removed a new paragraph here, ok?
$\langle \alpha,~g,~u\rangle$,
where $\alpha$ is a parameterised action, $\alpha \in\actions{\variables}$; 
$g\in\boolexprs{\variables}$ is a guard over variables of the source state and the 
action, and $u\in\assigns{\variables}$
assigns updated value for variables in the destination state. 
%%% SHORTEN BY LUDO
%If 
%$s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to $ then 
%% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
%		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
%		$\vars(g)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
%		$\vars(\codom(u))\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
%		\vars(\dom(u))\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$. 
%\noteLH{is dom and codom clear?}
%\noteSB{Not sure: I understand, because I know what is meant, but this is not a canonical use, so should be explained. Will be solved if we roll back to \mbox{$(?x_i := e_i)^{i \in I}$} or \mbox{$\overline{?x := e}$}.}
%\noteLH{COORDINATION: I believe we should remove the constraints}		
%\end{itemize}
\end{definition}
%Now we define
%pNets as constructors for hierarchical behavioural structures.
A pNet composes several pNets, pLTSs, and holes.
A  pNet  exposes
 global actions resulting from the synchronisation of internal actions in some sub-pNets. 
This synchronisation  is specified by  \emph{synchronisation vectors}.
Actions involved in the synchronisation vectors do
not need to distinguish input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.

@@

The red dotted rule expresses an implication, stating that if actions of the holes perform the designed actions and the condition is verified, then the variables are modified and the state changes. This implication however occurs in a restricted logic with predefined syntax, at a different logical level compared to  a classical
inference rule expressed in the logic of a scientific paper. 
 The open transition is  a logical implication that uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a general expressive, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.

@@

It requires the hole at label $B$ to fire a $fail(b)$ action, with the condition $b=true$ and the state variable $t<Max$. In this case, the global pNet goes from state $\ostate{11}$ that has an internal variable $t$ to state $\ostate{14}$ where the state variable $t'$ has the value $t+1$. This emits a global action $\alpha(t)$.

>>>>

\newcommand{\newCoord}[1]{{\color{blue}#1}}

@@

%\paragraph{Definitions.}
%\label{section:pNets}
A pLTS is a labelled transition system with variables occurring inside states,  actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
Variables of each state are pairwise disjoint.
\newCoord{The parameters of actions that appear on pLTS transitions are either input variables, bound when the action occurs and accepting any value (of the right type), or expressions using the value of other variables.}
We  define the set of parameterised actions a pLTS can use ($a$
ranges over action labels):
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
e_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $e_j^{j \in J}$ 
are expressions.

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
%\begin{itemize}
%\item
$S$ is a set of states; %$\vars(s)$ denotes the set of associated variables;
%\item
$s_0 \in S$ is the initial state;
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
%\item 
$\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
% SB: I removed a new paragraph here, ok?
$\langle \alpha,~g,~u\rangle$,
where $\alpha$ is a parameterised action, $\alpha \in\actions{\variables}$; 
$g\in\boolexprs{\variables}$ is a guard over variables of the source state and the 
action, and $u\in\assigns{\variables}$
assigns updated value for variables in the destination state. 
%%% SHORTEN BY LUDO
%If 
%$s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to $ then 
%% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
%		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
%		$\vars(g)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
%		$\vars(\codom(u))\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
%		\vars(\dom(u))\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$. 
%\noteLH{is dom and codom clear?}
%\noteSB{Not sure: I understand, because I know what is meant, but this is not a canonical use, so should be explained. Will be solved if we roll back to \mbox{$(?x_i := e_i)^{i \in I}$} or \mbox{$\overline{?x := e}$}.}
%\noteLH{COORDINATION: I believe we should remove the constraints}		
%\end{itemize}
\end{definition}
%Now we define
%pNets as constructors for hierarchical behavioural structures.
A pNet composes several pNets, pLTSs, and holes.
A  pNet  exposes
 global actions resulting from the synchronisation of internal actions in some sub-pNets, and some actions of the holes. \newCoord{As holes are process parameters, synchronisation with a hole has an obvious concrete meaning when a process is put inside the hole and emits the action. We  also define a semantics for open pNets with holes where open transitions express the fact that a pNet can performs a transition provided one or several holes emit some actions.}
\newCoord{This synchronisation  is specified by  \emph{synchronisation vectors} expressing the synchronous interaction between actions inside sub-pNets and holes, data transmission is expressed classically using action parameters.}
Actions involved in the synchronisation vectors do
not need to distinguish input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.

@@

\newCoord{The red dotted rule expresses an implication, stating that if actions of the holes perform the designed actions and the condition is verified, then the variables are modified and the state changes. This implication however occurs in a restricted logic with predefined syntax, at a different logical level compared to  a classical
inference rule expressed in the logic of a scientific paper. 
 The open transition is  a logical implication that uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a general expressive, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.}

@@

It requires the hole at label $B$ to fire a $fail(b)$ action, with the condition $b=true$ and the state variable $t<Max$. In this case, the global pNet goes from state $\ostate{11}$ that has an internal variable $t$ to state $\ostate{14}$ where the state variable $t'$ has the value $t+1$. This emits a global action $\alpha(t)$. \newCoord{Figure~\ref{fig:ArchFailure:OA} illustrates a whole open automaton.}
